---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# zendown

<!-- badges: start -->

[![R-CMD-check](https://github.com/rfsaldanha/zendown/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/rfsaldanha/zendown/actions/workflows/R-CMD-check.yaml)

<!-- badges: end -->

A simple R package to download files from Zenodo.

## Installation

You can install the development version of zendown from [GitHub](https://github.com/) with:

``` r
# install.packages("remotes")
remotes::install_github("rfsaldanha/zendown")
```

## Example

First, you need to find the Zenodo deposition code. It is the number that appears on the end of a Zenodo deposition link. The code number also appears on the Zenodo DOI of the version.

```         
https://zenodo.org/records/10950327
```

Deposition code: 10950327

With the deposition code, get a list of files available at the deposition

```{r}
library(zendown)

res <- file_list(deposit_id = 10950327)
res
```

You can download all files.

``` r
zendown(file_list = res, dest = tempdir())
```

Or a single file by specifying its name.

``` r
zendown(file_list = res, file_name = "indi_0002_regsaude_res_year.parquet", dest = tempdir())
```

## zen4R

To fully explore the Zenodo API possibilities, check the [zen4R](https://github.com/eblondel/zen4R) package.
