---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# zendown

<!-- badges: start -->

[![R-CMD-check](https://github.com/rfsaldanha/zendown/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/rfsaldanha/zendown/actions/workflows/R-CMD-check.yaml)

<!-- badges: end -->

Access, download and locally cache files deposited on Zenodo easily.

## Installation

You can install the development version of zendown from [GitHub](https://github.com/) with:

``` r
# install.packages("remotes")
remotes::install_github("rfsaldanha/zendown")
```

## Example

```{r}
library(zendown)
```

This [Zenodo deposit](https://zenodo.org/records/10959197) contains RDS files from examples datasets.

First, you need to find the Zenodo deposit code. It is the number that appears on the end of a Zenodo deposit link. The code number also appears on the Zenodo DOI.

```         
https://zenodo.org/records/10959197
```

Deposition code: 10959197

With the deposit code and the desired file name, you can just access the file with the `zen_file` function.

```{r, echo=FALSE}
delete_mirror(deposit_id = 10959197)
```

```{r}
readRDS(file = zen_file(deposit_id = 10959197, file_name = "iris.rds")) |>
  head()
```

The function will create a local cache, avoiding re-downloading the file when you access the file again.

## zen4R

To explore the Zenodo API possibilities to problematically store files and other procedures, check the [zen4R](https://github.com/eblondel/zen4R) package.
