---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# zendown

<!-- badges: start -->

[![R-CMD-check](https://github.com/rfsaldanha/zendown/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/rfsaldanha/zendown/actions/workflows/R-CMD-check.yaml)
[![CRAN status](https://www.r-pkg.org/badges/version/zendown)](https://CRAN.R-project.org/package=zendown)
<!-- badges: end -->

Access, download and locally cache files deposited on Zenodo easily.

## Installation

You can install the stable release from CRAN.

``` r
install.packages("zendown")
library(zendown)
```
Or install the development version of from GitHub:

``` r
# install.packages("remotes")
remotes::install_github("rfsaldanha/zendown")
library(zendown)
```

## Example

```{r}
library(zendown)
```

This [Zenodo deposit](https://zenodo.org/records/10959197) contains RDS files from examples datasets.

First, you need to find the Zenodo deposit code. It is the number that appears on the end of a Zenodo deposit link. The code number also appears on the Zenodo DOI.

```         
https://zenodo.org/records/10959197
```

Deposition code: **10959197**

With the deposit code and the desired file name, you can just access the file with the `zen_file` function.

```{r}
my_iris <- zen_file(deposit_id = 10959197, file_name = "iris.rds") |>
  readRDS()

head(my_iris)
```

The function will create a cache on your machine with all accessed files, avoiding re-downloading them when you access some file again.

## Cache

By default, the cache is stored on a temporary folder that is cleaned when the R session is ended.

To use a persistent cache and other options, [check this article](https://rfsaldanha.github.io/zendown/articles/cache_options.html).

## More examples

```{r}
# https://zenodo.org/records/10848
zen_file(10848, "DOAJ_Soc_Licenses_Correct.csv") |>
  read.csv2() |>
  tibble::tibble() |>
  head()
```

```{r}
# https://zenodo.org/records/10947952
zen_file(10947952, "2m_temperature_max.parquet") |>
  arrow::read_parquet() |>
  tibble::tibble() |>
  head()
```

```{r}
# https://zenodo.org/records/10889682
zen_file(10889682, "total_precipitation_2023-09-01_2023-09-30_day_sum.nc") |>
  terra::rast() |>
  terra::plot(1)
```

## zen4R

To explore the Zenodo API possibilities to problematically store files and other procedures, check the [zen4R](https://github.com/eblondel/zen4R) package.
