---
title: "Cache options"
---

The `zendown` package functions provides some options to cache downloads from Zenodo, avoiding unnecessary downloads.

## Temporary cache

By default, all downloads are cached in a **temporary** folder linked to the R session. When the R session if finished (for example, restarting R or RStudio), the cache will be emptied.

## Persistent cache

A persistent cache can be useful, specially if you handle large files. With a persistent cache, the downloaded files will be available between R sessions.

## Changing defaults

You can opt to **always** use a persistent cache. For this, you need to create an R environment variable.

First, edit you `.Renviron` file with `usethis::edit_r_environ()`

At the end of the file, add this line:

```         
zendown_cache_type = "persistent"
```

Then, save the file and restart R. From now on, the cached files will be persistent in a system folder.

## Where are cache files?

### Temporary cache

The temporary cache files are stored at the temporary files folder of your system. To retrieve this folder path, you can use the following command:

```{r, eval=FALSE}
fs::path_temp("zendown")
```

Remember that this folder is temporary and will be deleted when you close or restart the R session.

### Persistent cache

When you use a persistent option, the cache files are stored at the default package cache folder by default. To retrieve this folder path, you can use the following command:

```{r, eval=FALSE}
tools::R_user_dir("zendown", which = "cache")
```

## User defined persistent cache folder

You also may also want to define a specific folder to be used as cache, like a external drive. For this, set another environment variable.

Edit you `.Renviron` file with `usethis::edit_r_environ()` and add the following line at the end:

```         
zendown_cache_dir = C:\my_dir\
```

Substitute \`C:\my\_dir\` with you cache folder path. Save the file and restart. From now on, the cache files will be stored on this folder.

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(zendown)
```

## Cache management

A large number of cached files can easily drawn the storage capacity, specially with persistent cache.

Use the `cache_report()` function to track the number of Zenodo cached deposits, number of files and storage usage.

Use the `delete_mirror()` function to delete a cached Zenodo deposit.
